---
title: "GSS challange"
author: "Candice Fallon"
date: "Spring 2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
GSS <-read.csv(file="GSScsv.csv",header=TRUE)
```
```{r}
GSS$RACE <- as.factor(GSS$RACE)
GSS$BIGBANG <- as.factor(GSS$BIGBANG)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is belief in the bigbang theory (BIGBANG) independent of race (RACE)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

```{r}
dd2 <- GSS %>% group_by(RACE,BIGBANG) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=RACE,y=count,fill=BIGBANG))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=RACE,y=prcnt*100,fill=BIGBANG)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

What I see in this plot, is that the majority of Black people so not believe in the BIGBANG theory. While in the RACE "Other", the majority does believe in the BIGBANG theory. White people have a 50/50 belief in the theory. Each RACE has a different attitude when it comes to the theory.

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~RACE + BIGBANG, data=GSS)
rowPerc(table1)
colPerc(table1)
```

```{r}
table2 <- xtabs(~RACE + BIGBANG, data=GSS)
rowPerc(table2)
colPerc(table2)
```

My observation of the numerical results are 80% of the Black population do not believe in the BIGBANG theory. 55% of the Other race do believe in the BIGBANG theory. White people have a 50/50 belief of the theory. There is a big difference between Black peoples belief compared to the rest of the groups belief in the BIGBANG theory, while there is not that much difference between the Others and Whites belief in the theory.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the inferential results, we can see that there is a significant difference between the P-value in the Pearson's Chi-squared test which is p-value = 0.01098. While in the Fisher's Exact Test for Count Data the p-value = 0.00905. The Fisher exact test tells me that theres is something significant going on, but I still don't know what that is.


#2X2 TEST
```{r}
chisq.test(table2)
chisqtestGC(table2)
```



## A function to run multiple fisher tests
### Upload the fisherXY.R function then source it

```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```
```{r}
fisherXY(GSS$RACE,GSS$BIGBANG,printAll = T)
```


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

In conclusion, the results of the 2x2 test was ran to see each of the RACE's beliefs on the BIGBANG theory and you can see that there is a significant difference between the Blacks and the other RACES.

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
